Redundant calibration as a complex optimization problem

### 2017-01-06 paper.tex
added some more structure to paper

\section{Preliminaries}

\subsection{Conjugate Gradient Method}
\label{sec:conj_grad}
The Conjugate Gradient (CG) method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
\begin{equation}
\label{eq:linear_system}
\bb = \bA\bx,
\end{equation}
where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.  
Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
The iterative CG method is presented in Algorithm \ref{algo:CG}.

@@

\begin{algorithm}
\caption{Conjugate Gradient Method. Inputs: $\bA$, $\bb$, a starting value $x$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}. \citep{Shewchuk1994}.}\label{algo:CG}
\begin{algorithmic}[1]
\State $k \gets 0$
\State $\br \gets \bb - \bA\bx$
\State $\bp \gets \br$
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\delta_0 \gets \delta_{\textrm{new}}$
\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
\State $\alpha_k \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
\State $\bx \gets \bx + \alpha_k\bp$
\State $\br \gets \br - \alpha\bA\bp$
\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\beta_{k} \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
\State $\bp \gets \br + \beta\bp$
\State $k \gets k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

\subsubsection{Computational Complexity}
\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
The mathematical definition of $\kappa(\bA)$ is
\begin{equation}
\label{eq:kappa}
\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
\end{equation}
where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
The computational complexity of CG is therefore 
\begin{equation}
\label{eq:cg_bound}
\mathbb{O}(\sqrt{\kappa}m). 
\end{equation}
It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems, since in those cases $m < P^2$.

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ (as the largest and smallest eigenvalue of the Identity matrix is equal to one) becoming unity.
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
\begin{equation}
\bM = \bA\odot\bI, 
\end{equation}
where $\bI$ is the identity matrix. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.

\subsubsection{Normal Equation}
\label{sec:normal}
Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
\begin{equation}
 \bb = \bB\bx,
\end{equation}
where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
\begin{equation}
\label{eq:normal_equation}
\bB^H\bB\bx = \bB^H\bb 
\end{equation}
to estimate the $\bx$ that minimizes
\begin{equation}
\label{eq:norm_ls}
\|\bb-\bB\bx\|_F^2. 
\end{equation}
There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
\begin{enumerate}
\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
\begin{equation}
\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
\end{equation}
\end{enumerate}

Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. The CG method actually has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}.

In light of \citet{Lu2015}, the two observations in the numbered list we presented earlier in this section imply that Eq.~\eqref{eq:normal_equation}
can be solved by employing the CG method. The solution returned by CG is however only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
obtained by CG, whether it be unique or not, does however minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.

@@

\subsection{Least Squares}
\label{sec:ls}
The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
\begin{equation}
\label{eq:least_squares}
\min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
\end{equation}
where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model data vector and $\bzeta$ is the model parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
The main objective of least squares is to find the model parameters which minimizes the squared difference between the observed data and the model predicted data. Moreover, $\|*\|_F$ denotes the 
Frobenius norm.

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda_{\textrm{LM}}\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda_{\textrm{LM}}$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. These update rules are used in an iterative manner to 
determine new estimates of $\bzeta$.

The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods. The reason being,
many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
classical definition of differentiation is used then $\frac{\partial z}{\partial \conj{z}}$, where $z \in \mathbb{C}$, does not exist.

@@

\section{Problem Statement}

\subsection{Scalar Measurement Equation}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We try to estimate the value of
$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.

@@

\subsection{Indexing Functions}

In general there are two ways of indexing the true visibilities associated with different baselines. We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
\begin{equation}
12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
\end{equation}
We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.

@@

\subsection{Redundant Measurement Equation}
If the array is redundant and the function $\phi_{pq}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
\begin{equation}
\label{eq:vis_red}
d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq}.
\end{equation}
Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
\begin{equation}
\label{eq:vis_red_vec}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{align}
 \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
 \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
\end{align}
The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).

We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
\begin{equation}
\label{eq:least_squares_red}
\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
\end{equation}
where
\begin{align}
 \by &= [y_1,\cdots,y_L]^T, & \bg &=[g_1,\cdots,g_N]^T,\nonumber\\
 \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
 \end{align}
The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
We plot the number of unique baseline groups $L$ and the number of parameters we need to estimate $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
\begin{equation}
L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}, 
\end{equation}
and
\begin{equation}
L = N - 1. 
\end{equation}

\begin{figure*}
\centering
\subfigure[$L$.]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
\subfigure[$P$.]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
\caption{In the left panel we plot the number of unique baseline groups $L$ one can form when using a hexagonal (red curve) and regular east-west layout (blue curve) as 
a function of the number of antennas $N$ in the array. In the right panel we plot the total number of parameters we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as 
a function of the number of antennas $N$ in the array. 
\label{fig:pl}} 
\end{figure*}

\subsection{Signal-to-noise Ratio}
The following SNR (signal-to-noise ratio) definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

HERE I CAN WRITE SOMETHING ABOUT THE LIMITATIONS OF THE PAPER.

@@

>>>>

\section{Preliminaries}
There are two major pre-existing algorithms we use throughout this paper, namely the Conjugate Gradient (CG) method and least-squares parameter estimation.
We respectively discuss the CG method and least-squares in Section~\ref{sec:conj_grad} and Section~\ref{sec:ls}.

\subsection{Conjugate Gradient Method}
\label{sec:conj_grad}
The CG method was independently discovered by \citet{Hestenes1973} and \citet{Stiefel1952}. They later published a joint paper, which is now considered as the seminal
reference on CG \citep{Hestenes1952}. The iterative CG method is used to solve a particular class of linear system. The CG method is generally used to solve
\begin{equation}
\label{eq:linear_system}
\bb = \bA\bx,
\end{equation}
where $\bb\in\mathbb{C}^P$ is a known complex column vector of size $P$, $\bx\in\mathbb{C}^P$ is an unknown complex column vector with the same length as $\bb$, and $\bA\in\mathbb{C}^{P\times P}$ is a square positive-definite Hermitian matrix with the appropriate dimensions.  
Moreover, using the CG method to solve Eq.~\eqref{eq:linear_system} will only be computationally advantageous if $\bA$ is also sparse (contain many zero entries).
Using the iterative CG method to solve large sparse linear sytems was popularized by \citet{Reid1971}. A good tutorial on the CG method can be found in \citep{Shewchuk1994}.
The iterative CG method is presented in Algorithm \ref{algo:CG}.

@@

\begin{algorithm}
\caption{Conjugate Gradient Method. Inputs: $\bA$, $\bb$, a starting value for $\bx$, maximum number of iterations $k_{\textrm{max}}$ and an error tolerance $\epsilon<1$ Output: $\bx$ the solution to Eq.~\eqref{eq:linear_system}. \citep{Shewchuk1994}.}\label{algo:CG}
\begin{algorithmic}[1]
\State $k \gets 0$
\State $\br \gets \bb - \bA\bx$
\State $\bp \gets \br$
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\delta_0 \gets \delta_{\textrm{new}}$
\While{$k\leq k_{\textrm{max}}$ and $\delta_{\textrm{new}} > \epsilon^2\delta_0$}
\State $\alpha_k \gets \frac{\delta_{\textrm{new}}}{\bp^H\bA\bp}$
\State $\bx \gets \bx + \alpha_k\bp$
\State $\br \gets \br - \alpha\bA\bp$
\State $\delta_{\textrm{old}} \gets \delta_{\textrm{new}}$ 
\State $\delta_{\textrm{new}} \gets \br^H\br$
\State $\beta_{k} \gets \frac{\delta_{\textrm{new}}}{\delta_{\textrm{old}}}$
\State $\bp \gets \br + \beta\bp$
\State $k \gets k + 1$
\EndWhile
\end{algorithmic}
\end{algorithm}

\subsubsection{Computational Complexity}
\citet{Kaniel1966} was able to determine convergence bounds for CG using Chebyshev polynomials. A much more in depth analysis of CG convergence can be found in
\citep{Sluis1986}. As can be seen from Algorithm \ref{algo:CG}, the CG method makes use of a major outer loop (line 6 of Algorithm \ref{algo:CG}). It turns out that at worst 
the number of major iterations that CG requires to converge\footnote{If the $\bA$-norm is used to calculate the error.} is proportional to $\sqrt{\kappa}$, where $\kappa$ is the spectral condition number of the matrix $\bA$.
The mathematical definition of $\kappa(\bA)$ is
\begin{equation}
\label{eq:kappa}
\kappa(\bA) = \frac{\lambda_{\textrm{max}}}{\lambda_{\textrm{min}}}, 
\end{equation}
where $\lambda_{\textrm{max}}$ and $\lambda_{\textrm{min}}$ respectively denote the largest and the smallest eigenvalue of $\bA$.
Moreover, the most expensive operation that is conducted within each major loop of Algorithm~\ref{algo:CG} is the vector-matrix product $\bA\bp$, which has a complexity of $\mathbb{O}(m)$, where $m$ is the number of non-zero entries in $\bA$.
The computational complexity of CG is therefore 
\begin{equation}
\label{eq:cg_bound}
\mathbb{O}(\sqrt{\kappa}m). 
\end{equation}
It should now also be clear to the reader why the CG method is especially applicable to sparse linear systems. In sparse linear systems $m < P^2$.

\subsubsection{Preconditioning}
\label{sec:precon}
Preconditioning is a technique which allows us to improve the condition number of a matrix. Let $\bM$ be a positive-definite Hermitian matrix.
We can now indirectly solve for $\bx$ by using 
\begin{equation}
\label{eq:prec}
\bM^{-1}\bA\bx = \bM^{-1}\bb.
\end{equation}
If $\bM$ is a good preconditioner then $\kappa(\bM^{-1}\bA) \ll \kappa(\bA)$, i.e. a good preconditioner lowers the condition number of a matrix which in turn increases the convergence speed 
of CG (see Eq.~\eqref{eq:cg_bound}). The best possible choice of $\bM$ would therefore be $\bA$, which would lead to $\kappa$ becoming unity (as the largest and smallest eigenvalue of the Identity matrix is equal to one).
If $\kappa$ is one then Eq.~\eqref{eq:cg_bound} simplifies and becomes $\mathbb{O}(m)$. However, if we knew the inverse of $\bA$ then obtaining $\bx$ would be trivial and there would be no need for the CG method in the first place. 
In short, a good preconditioner $\bM$ must approximate $\bA$ fairly accurately. However, if it is not much less expensive to 
compute the inverse of $\bM$ than that of $\bA$ we actually gain nothing (see Eq.~\eqref{eq:prec}).

If the diagonal entries of $\bA$ is much more significant than its off-diagonal entries then the 
Jacobian preconditioner is a good choice of $\bM$. The Jacobian preconditioner is computed as follows:
\begin{equation}
\bM = \bA\odot\bI, 
\end{equation}
where $\bI$ is the identity matrix and $\odot$ denotes the Hadamard product. Since the Jacobian preconditioner is a diagonal matrix its inverse can be obtained trivially.

\subsubsection{Normal Equation}
\label{sec:normal}
Lets assume that instead of Eq.~\eqref{eq:linear_system} we have a more general linear system, i.e.
\begin{equation}
 \bb = \bB\bx,
\end{equation}
where $\bb\in\mathbb{C}^K$, $\bx\in\mathbb{C}^P$  and $\bB\in\mathbb{C}^{K \times P}$, with $K > P$. We can use the normal equation 
\begin{equation}
\label{eq:normal_equation}
\bB^H\bB\bx = \bB^H\bb, 
\end{equation}
to estimate the $\bx$ that minimizes
\begin{equation}
\label{eq:norm_ls}
\|\bb-\bB\bx\|_F^2. 
\end{equation}
In this paper, we denote the Hermitian transpose with $(*)^H$ and the Frobenius norm with $\|*\|_F$.

There are a few important observations that we can make while inspecting Eq.~\eqref{eq:normal_equation}:
\begin{enumerate}
\item $\bB^H\bB$ is a square positive semi-definite Hermitian matrix by construction.
\item $\bB^H\bb$ is in the column range of $\bB^H\bB$; this follows trivially from 
\begin{equation}
\bB^H\bb \in \mathcal{R}(\bB^H) \implies \bB^H\bb \in \mathcal{R}(\bB^H\bB).   
\end{equation}
\end{enumerate}
In this paper, we denote the column range of a matrix $\bA$ with $\mathcal{R}(\bA)$.

Recall that we mentioned that in general we can only solve Eq.~\eqref{eq:linear_system} with CG if $\bA$ is a square positive-definite Hermitian matrix. The CG method actually has an even broader applicability, it can be used to solve Eq.~\eqref{eq:linear_system} even if $\bA$ is an Hermitian positive semi-definite
square matrix as long as $\bb$ is in the column range of $\bA$ \citep{Lu2015}. In light of \citet{Lu2015}, the two observations in the numbered list we presented in the previous paragraph imply that Eq.~\eqref{eq:normal_equation}
can be solved by employing the CG method. The solution returned by CG is however only unique if $\bB^H\bB$ is positive-definite (invertible). The solution
obtained by CG, whether it be unique or not, does however minimize Eq.~\eqref{eq:norm_ls} in a least squares sense.

@@

\subsection{Least Squares}
\label{sec:ls}
The method of nonlinear least squares is tanatamount to solving the following optimization problem:  
\begin{equation}
\label{eq:least_squares}
\min_{\bzeta} \Lambda(\bzeta) = \min_{\bzeta} \|\brho\|_F^2 = \min_{\bzeta} \|\bdelta - \bnu(\bzeta)\|_F^2, 
\end{equation}
where $\Lambda$ is known as the objective function, $\brho$ is the residual vector, $\bdelta$ is the observed data vector, $\bnu$ is the model-predicted data vector and $\bzeta$ is the model-parameter vector that minimizes Eq.~\eqref{eq:least_squares}.
In Eq.~\eqref{eq:least_squares}, we assume that the optimization problem is real valued.
The main objective of least squares is to find the model-parameters which minimizes the squared difference between the observed data and the model-predicted data.

The most standard way of solving Eq.~\eqref{eq:least_squares} is to use a gradient type of minimization algorithm. In this paper we will be focusing on the following gradient minimization algorithms: Gauss-Newton and Levenberg-Marquardt \citep{Levenberg1944,Marquardt1963}. 
The GN update step is defined as
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ)^{-1} \bJ^T\brho,
\end{equation}
while the LM update takes the following form
\begin{equation}
 \Delta \bzeta = (\bJ^T\bJ+\lambda\bD)^{-1} \bJ^T\brho,
\end{equation}
where $\bJ$ is the Jacobian matrix (i.e. the derivative of the model $\bnu$ with respect to the parameter vector $\bzeta$), $\lambda$ denotes the 
damping factor, while $\bD$ is a diagonal matrix whose diagonal entries are equal to the diagonal entries of $\bJ^T\bJ$. In this paper we denote 
matrix transposition with $(*)^T$.
These update steps can be used in an iterative manner to determine new estimates of $\bzeta$.

The gradient based minimization algorithms that are generally used to solve least squares problems (i.e. GN and LM) generally require the model $\bnu$ to be differentiable
towards each model parameter. When the least squares problem is complex it becomes less straightforward to apply these gradient based minimization methods. The reason being,
many complex functions are not analytic (no Taylor expansion around a fiducial point exists) and therefore not differentiable if the classic notion of differentiation is used. For instance if the 
classical definition of differentiation is used then $\frac{\partial \conj{z}}{\partial z}$, where $z \in \mathbb{C}$, does not exist.

@@

\section{Problem Statement}
In Section~\ref{sec:s_mes}, we present the scalar measurement equation. Next, we define some indexing functions in Section~\ref{sec:indexing}. The measurement 
equation and the indexing functions from Section~\ref{sec:indexing} are then used in Section~\ref{sec:r_mes} to phrase redundant calibration as a least-squares problem, i.e. Eq.~\eqref{eq:least_squares_red}.
We present the Signal-to-noise ratio definition we used in this paper in Section\ref{sec:snr}. The general scope and limitations of this paper is discussed in Section~\ref{sec:scope}.

\subsection{Scalar Measurement Equation}
\label{sec:s_mes}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the scalar measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are unknown. We try to estimate the value of
$g_p$, $g_q$ and $y_{pq}$ by calibrating our data and then imaging the result.

@@

\subsection{Indexing Functions}
\label{sec:indexing}
In general there are two ways of indexing the true visibilities associated with different baselines. We can use the composite index $pq$, which is the strategy that is followed in Eq.~\eqref{eq:vis_definition},
or we can use a single unique baseline index which we assign in numerical order to each composite index; assuming the composite indexes are arranged in standard baseline order.
The composite indexes $pq$ are in standard baseline order when they are arranged as follow: $12, 13, 1N, 23\cdots$. Let 
\begin{equation}
\alpha_{pq} =
\begin{cases}
(q-p) + (p-1)\left (N-\frac{1}{2}p \right ) & \textrm{if}~p<q\\
%B + (p-q) + (q-1)\left (N-\frac{1}{2}q \right )) & \textrm{if}~p>q\\
%B+q + \frac{1}{2}(p-1)(p-2) & \textrm{if}~p>q\\
0 & \textrm{otherwise}
\end{cases}.
\end{equation}
We can use $\alpha_{pq}$ to map composite indexes to unique baseline indexes, i.e. 
\begin{equation}
12,13,\cdots,1N,23,\cdots \xrightarrow{\alpha_{pq}} 1,2,3,\cdots,B.
\end{equation}
We denote the number of baselines in the array with $B$, i.e. $B = \frac{N^2-N}{2}$.

@@

\subsection{Redundant Measurement Equation}
\label{sec:r_mes}
If the array is redundant and the function $\phi_{pq}$ is known we can rewrite Eq.~\eqref{eq:vis_definition} as
\begin{equation}
\label{eq:vis_red}
d_{pq} = g_{p}\conj{g_q}y_{\phi_{pq}} + n_{pq}.
\end{equation}
Eq.~\eqref{eq:vis_red} can also be expressed in the following vector form 
\begin{equation}
\label{eq:vis_red_vec}
\bd = \bv + \bn, 
\end{equation}
where 
\begin{align}
 \left [ \bd \right]_{\alpha_{pq}} &= d_{pq}, & \left [ \bv \right ]_{\alpha_{pq}} &= g_p y_{\phi_{pq}} \conj{g_q},\nonumber\\
 \left [ \bn \right ]_{\alpha_{pq}} &= n_{pq}. &  &\label{eq:vec_definitions}
\end{align}
The vectors in Eq.~\eqref{eq:vec_definitions} are column vectors of size $B$ (i.e. $p<q$).

We have now defined enough notation to enable us to formulate redundant calibration as a least-squares problem:
\begin{equation}
\label{eq:least_squares_red}
\min_{\bz} \Lambda(\bz) = \min_{\bz} \|\br\|_F^2 = \min_{\bz} \|\bd - \bv(\bz)\|_F^2, 
\end{equation}
where
\begin{align}
 \by &= [y_1,\cdots,y_L]^T, & \bg &=[g_1,\cdots,g_N]^T,\nonumber\\
 \bz &= [\bg^T,\by^T]^T. &  &\label{eq:parm_definitions}
 \end{align}
The number of model parameters that we need to solve for in Eq.~\eqref{eq:least_squares_red} is equal to $P = 2(N+L)$, since redundant calibration is a complex problem.
We plot $L$ and $P$ for two different geometric layouts in Fig~\ref{fig:pl}.
The analytic expressions for the red and the blue curve in Fig.~\ref{fig:ln} are respectively equal to
\begin{equation}
L = 2N-\frac{1}{2}\sqrt{12N-3}-\frac{1}{2}, 
\end{equation}
and
\begin{equation}
L = N - 1. 
\end{equation}

\begin{figure*}
\centering
\subfigure[$L$]{\includegraphics[width=0.47\textwidth]{./ln.pdf}\label{fig:ln}}
% V_R_3.pdf: 585x441 pixel, 72dpi, 20.64x15.56 cm, bb=0 0 585 441
\subfigure[$P$]{\includegraphics[width=0.47\textwidth]{./pn.pdf}\label{fig:pn}}
\caption{In the left panel we plot the number of unique baseline groups $L$ one can form when using a hexagonal (red curve) and regular east-west layout (blue curve) as 
a function of the number of antennas $N$ in the array. In the right panel we plot the total number of parameters $P$ we need to estimate when employing redundant calibration for a hexagonal (red curve) and regular east-west layout (blue curve) as 
a function of the number of antennas $N$ in the array. 
\label{fig:pl}} 
\end{figure*}

\subsection{Signal-to-noise Ratio}
\label{sec:snr}
The following SNR (signal-to-noise ratio) definition is used in this paper  
\begin{equation}
\label{}
\textrm{SNR} = \frac{<\bv\odot\conj{\bv}>_{\nu,t,pq}}{<\bn\odot\conj{\bn}>_{\nu,t,pq}}, 
\end{equation}
where $<*>_{\nu,t,pq}$ denotes averaging over frequency, time and baselines and $\odot$ denotes the Hadamard product. The definition we use here is based on the SNR definitions used in \citet{Liu2010} and \citet{Marthi2014}.

\subsection{Paper Scope}
\label{sec:scope}
We only deal with per-timeslot and frequency channel calibration. No solution smoothing  
is applied between the different timeslots and channels. The reader interested in this is referred to \citep{Zheng2014}. 
Moreover, we do not explicitly deal with any of the degeneracies normally associated with redundant calibration \citep{Zheng2014,Kurien2016}. Similar to the approach taken in \citep{Marthi2014}, we assume that the degeneracies present in the redundant calibration solutions are removed with a final round of skymodel-based calibration.

@@

\section{Conclusion}
Hallo
